Diagnosing the sensitivity of grounding line flux to changes in sub-ice shelf melting

### 2019-07-24 draft_v2/draft_v2.tex
Updates from Overleaf

\usepackage{appendix}

@@

\begin{document}
\title{Diagnosising the sensitivity of grounding line flux to changes in sub-ice shelf melting}

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{./figs/mismip_larsenc.pdf}
\caption{Plan view of steady-state surface speed for a) MISMIP+ and b) Larsen C Ice Shelf. The black curves show the grounding lines.}
\label{mismip_larsenc}
\end{figure}

@@

Our experiments are conducted in a manner similar to those of \citet{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, we use a uniform, hexagon-type, around 2-km mesh (for simplicity we use the term 2-km in the following). For the Larsen C Ice Shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we perturb the thickness of the cell (hexagon) in the mesh. For Larsen C, we only use 20$\times$20 km square boxes \citep[the same resolution as in][]{reese2018}. Lastly, for the MISMIP+ 2-km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is laterally buttressed).  We also perturb only half of the ice shelf, $y>40$ km, due to the symmetry about the center line, though we analyze the response to these perturbations over the entire model domain. \textbf{TZ: I am afraid I still want to use Python to make Fig. 1 to show the coordinates}

Similarly to \citet{reese2018}, we define a GLF ``response number'',
\begin{equation}
N_r = \left(\frac{R}{P}\right)^k,
\end{equation}
where $R$ is the ice flux change integrated along the entire grounding line, $P$ is the mass change associated with a single grid-box perturbation and $k$ is a power-law index that allows for the possibility of a nonlinear relationship between ice shelf buttressing and the change in GLF \citep[see also][]{schoof2007}. Here, we use $k=1/n$ with $n=3$.

Despite the existence of many possible factors linking GLF to ice shelf properties (ice flow direction, horizontal gradients in ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc.), here we mainly examine model stress fields and the distance between perturbation points and the GL. This is because, as we show below, these factors correlate closely with the sensitivity of changes in GLF to the imposed ice shelf perturbations. \textbf{SP: Maybe we can skip the previous sentences and just skip to some form of the next one? We can motivate this simply by saying that the reason we use a buttressing number like Furst et al is that it's a popular/standard and easy way to connect local ice shelf perturbations to the changes in GLF from the Reese experiments.} To incorporate the local stress field and its buttressing capacity into our analysis, we also calculate a ``buttressing number'', $N_b$, analogous to that from \citet{furst2016},
\begin{equation}
N_b\left(\mathbf{n}\right)=1-\frac{\mathbf{n}\cdot\mathbf{\sigma}\cdot\mathbf{n}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=\frac{1}{2}\left(1-{\rho_i}/{\rho_w}\right)gH$) and $\rho_i$ (910\,kg\,m$^{-3}$) and $\rho_w$ (1028\,kg\,m$^{-3}$) are the density of ice and ocean water, respectively. $\sigma_{nn}=\mathbf{n}\cdot\mathbf{\sigma}\cdot\mathbf{n}$ is the normal stress along a specified horizontal direction given by the normal vector $\mathbf{n}$ (see details in below).

@@

A decrease in ice shelf buttressing tends to lead to an increase in GLF \citep[e.g.,][]{gagliardini2010} and intuitively we expect that the GLF should be relatively more sensitive to melt changes that occur in regions of relatively larger buttressing. Here, we aim to better understand and quantify the relationship between the local ice shelf buttressing ``strength'' (given by $N_b$) and changes in GLF (given by $N_r$). Can the change in GLF for a given melt perturbation be quantified (in the predictive sense) as a function of the buttressing number? In Fig.~\ref{mismip_Nb_GLF_regression}, we show correlations between $N_b$ and $N_r$ for perturbations applied to the MISMIP+ domain when $N_b$ from (\ref{butN}) is defined based on the direction of the first principal stress ($\sigma_{p1}$) \footnote{We expand on our reasons for choosing $\sigma_{p1}$ below.}. In Fig.~\ref{mismip_Nb_GLF_regression}b, we include all perturbed ice shelf locations in the comparison and find relatively weak $N_b$:$N_r$ correlations. In Fig.~\ref{mismip_Nb_GLF_regression}c, we remove from the comparison points that are 1) weakly buttressed ($x>480$\,km, where the ice shelf becomes unconfined) and 2) within 10 km of the GL, and a much stronger $N_b$:$N_r$ correlation emerges. In Fig.~\ref{mismip_Nb_GLF_regression}d, we plot the $N_b$:$N_r$ correlation for locations where $x>480$\,km and the magnitude of the thickness gradient $\left|\nabla H\right|<7\times 10^{-3}$ (a similar plot with the buttressing number defined based on the direction of the second principal stress is shown in Fig.~\ref{sigma_p2_example}). The strong similarities between Figures \ref{mismip_Nb_GLF_regression}c and \ref{mismip_Nb_GLF_regression}d suggest that the seemingly \textit{ad hoc} ``distance from the grounding line'' constraint is important for removing areas of complex geometry (and hence complex dynamics) from the $N_b$:$N_r$ comparisons. We discuss possible reasons for this further below.

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/mismip_Nb_GLF_regression.pdf}
    \caption{(a) MISMIP+ steady-state geometry. Color represents the magnitude of the ice thickness gradient and the white line represents the GL. (b) $N_b$:$N_r$ correlation for all perturbation points. (c) $N_b$:$N_r$ correlation for perturbation points within the confined region of the shelf ($x>480$) and $>$10 km from the GL; (d) $N_b$:$N_r$ correlation for perturbation points within the confined region of the shelf and with a thickness gradient magnitude $\left|\nabla H\right|<7$x$10^{-3}$. In (c) and (d), the orange dots in the insets represent the GL grid cells, and the blue dots the perturbation grid cells. \textbf{SP: Matt suggested expanding this to include 2 more panels using the points in (d) -- the flow direction and the p2 direction (the goal being to show that the correlations are pretty bad). Then we could also add another panel along the bottom which is the 'continuous' version of these (i.e., the red curve in Fig. 3a).}}
\label{mismip_Nb_GLF_regression}
\end{figure}

%\subsection{The buttressing directions}
%\subsection{Correlation dependence on buttressing direction}
\subsection{Directional dependence of buttressing}

The buttressing number, $N_b$, is computed using the normal stress ($\sigma_{nn}$) projected onto a specific direction, see Equation (\ref{butN}). Therefore, the buttressing number at any perturbation point varies depending on the chosen direction. \citet{furst2016} calculated $N_b$ along two directions, the ice flow direction ($\mathbf{n}_f$) and the direction corresponding to the second principal stress ($\sigma_{p2}$) ($\mathbf{n}_{p2}$). They found that the latter---the direction corresponding to the maximum compressive stress (or the least extensional stress)---best identified ``passive'' regions, where the ice shelf provides little or no buttressing. In Fig.~\ref{mismip_r2_all_direction}, we plot the correlation coefficients ($r^2$) between $N_r$ and $N_b$ where the direction corresponding to $\sigma_{nn}$ varies continuously between 0 and 180$^\circ$ (as a function of the angle $\Delta \phi$) relative to the direction corresponding to the first principle stress ($\sigma_{p1}$) ($\mathbf{n}_{p1}$). We find the largest correlation coefficient ($r^2>0.9$) when $N_b$ is aligned with $\mathbf{n}_{p1}$ ($\Delta \phi=0^\circ$) and the smallest correlation coefficient ($r^2<0.4$) when $N_b$ is aligned with $\mathbf{n}_{p2}$ ($\Delta \phi=90^\circ$). We can also see the corresponding variations in buttressing number $N_b$ in Fig.~\ref{Nb_Deltaphi}, where the maximum buttressing number is calculated from $N_r(\mathbf{n}_{p2}$) according to Equation (\ref{butN}).
% for different directions with respect to the first principal stress and flow direction, respectively (Fig \ref{mismip_r2_all_direction}). Differently, we find that $N_b$ along the $\sigma_{p1}$ direction shows the best regression performance, whereas the $\sigma_{p2}$ appears to show the weakest correlations to $N_r$. This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$).

@@

A similar conclusion can be reached when the buttressing number is calculated by projecting $\sigma_{nn}$ onto $\mathbf{n}_{f}$\footnote{In \citet{furst2016}, the ice flow direction is the second direction considered when calculating the local ice shelf buttressing number (with the first direction considered being $\mathbf{n}_{p2}$).}. In Fig.~\ref{mismip_r2_all_direction}a, the blue line shows the $N_b$:$N_r$ correlation for the case of rotating relative to $\mathbf{n}_{f}$. %(the flow direction experiments are analyzed independently from the $\sigma_{p1}$ direction experiments, and can therefore be used as an approach for verification). 
For around 50\% of the perturbation points, the angles of flow directions are 30--50$^\circ$ ($\Delta\varphi$) more than that of the directions of first principle stresses (from Fig.~\ref{mismip_r2_all_direction}b), which is consistent with the phase difference between the blue and red curves in Fig.~\ref{mismip_r2_all_direction}a. For example, for the $\mathbf{n}_{f}$ curve (blue), we have the smallest value at $\Delta\phi\approx$ 38$^\circ$, i.e., 38$^\circ$ relative to the flow direction. Since the angles of $\mathbf{n}_{f}$ are 30--50$^\circ$ more than that of $\mathbf{n}_{p1}$, the lowest value should occur at 68--88$^\circ$ ($\Delta\phi+\Delta\varphi$) relative to $\mathbf{n}_{p1}$, which is corresponding to $\mathbf{n}_{p2}$. This is in consistent with the position of the smallest value for the $\sigma_{p1}$ direction curve (red; $\Delta\phi\approx$ 90$^\circ$). %\textbf{(SFP: See comments in source text regarding this paragraphy.)}

@@

Based on the discussion above and the work of \citet{furst2016}, it appears that $N_b(\mathbf{n}_{p2})$ provides a good indicator for the maximum buttressing at any location on the shelf and for identifying passive ice shelf areas. Yet it may not be a good metric for quantifying the sensitivity of GLF response to sub-shelf basal melt perturbations. In the following sections, we explain how and why changes in GLF are sensitive to the choice of direction when calculating the ice shelf buttressing number.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/mismip_r2_all_direction.pdf}
    \caption{(a) $N_b$:$N_r$ correlation coefficients for $\sigma_{nn}$ values rotated counterclockwise by $\Delta\phi$ degrees relative to $\mathbf{n}_{p1}$ (red) and $\mathbf{n}_{f}$ (blue). (b) Histogram of the angular differences between $\mathbf{n}_{f}$ and $\mathbf{n}_{p1}$. The perturbation points analyzed here are those shown in the inset of Fig.~\ref{mismip_Nb_GLF_regression}d. \textbf{SP: Matt/Xylar suggested breaking the velocity part of this fig out into its own fig.}}
    \label{mismip_r2_all_direction}
\end{figure}

%\subsection{Possible controlling factors}
%\subsection{Understanding the dependence on buttressing direction}
%\subsection{Direction of velocity changes around perturbation points}
\subsection{Velocity changes in the vicinity of perturbation points}

The correlation between changes in GLF and normal stress (buttressing) directions can be partly understood by examining the spatial patterns of velocity change associated with thickness perturbations. %Further clues of the impacts from different chosen directions can be found in Fig.~\ref{perturb_theta_velchange}. 
In Fig.~\ref{perturb_theta_velchange}, we plot histograms of the maximum (red) and minimum (blue) speed changes around each perturbation point as a function of angular distance relative to $\mathbf{n}_{p1}$ ($\Delta\phi$, as in Fig.~\ref{mismip_r2_all_direction}a). That is, we count the number of neighboring cells that have maximum (most increased or least decreased) and minimum (most decreased or least increased) speed changes (for each cell, there is only one neighbor with maximum or minimum speed change).  %Figure \ref{perturb_theta_velchange} contains only the points from Fig.~\ref{mismip_Nb_GLF_regression}d (i.e., filtered according to ice thickness gradient). 
The maximum speed changes cluster within 0 to 45$^\circ$ of $\mathbf{n}_{p1}$ and the minimum speed changes cluster within 60 to 90$^\circ$, indicating that the majority of speed increases in the immediate region of a perturbation are more closely aligned with $\mathbf{n}_{p1}$ (and likewise, the majority of velocity decreases are more closely aligned with $\mathbf{n}_{p2}$). This finding supports the hypothesis that the thickness perturbations induce velocity increases along favored directions and that, for the MISMIP+ domain, $\mathbf{n}_{p1}$ is favored. 
\textbf{SP: I think it would help if we try to connect this more clearly to the next section. That is, here we show that a local perturbation leads to speed increases along preferred directions. Below, we show that local perturbations lead to similar speed increases (also along preferred directions) but at the GL.}

@@

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/perturb_theta_velchange.pdf}
    \caption{Histograms for the maximum (red) and minimum (blue) percent speed change in grid cells adjacent to a thickness perturbation point, plotted as a function of angular distance with respect to the $\sigma_{p1}$ direction. Points analyzed are those from Fig.~\ref{mismip_Nb_GLF_regression}d.}
	\label{perturb_theta_velchange}
\end{figure}

%\subsection{The changes of velocity and stress along GL}
\subsection{Speed and stress changes along the GL}

Further insight into the importance of $\mathbf{n}_{p1}$ is given by examining changes in the state of stress and velocity as a result of a perturbation in local ice thickness. To this end, we define the metric $\Upsilon$,  
\begin{equation}
    \Upsilon= \textrm{Corr}\left(\frac{\sigma_{p}-\sigma_{c}}{\sigma_{c}}, \frac{u_{p}-u_{c}}{u_{c}}\right),
\label{upsilon}
\end{equation}
%\begin{equation}
%    \Psi= \frac{\sigma_{p}-\sigma_{c}}{\sigma_{c}}-\frac{u_{p}-u_{c}}{u_{c}},
%\label{psi}
%\end{equation}
\textbf{SP: need to update this equation to clarify the following things: 1) upsilon is scalar function of x,y (that is, we calculate a single value of it for each perturbation point on the ice shelf), 2) the two terms in parentheses are vectors of points along the GL, 3) the operation here is a dot product, so that the vectors collapse to a scaler value, which is what is evalutated).}
where the subscripts $p$ and $c$ denote the perturbation experiments and the ``control'' (i.e., the initial condition), respectively, and $\sigma$ and $u$ denote the stress component and ice speed along the GL, respectively. $\Upsilon$, a correlation coefficient, is a measure of the consistency between the magnitude and sign of the change in the model stress and velocity states along the GL between the control and perturbation experiments. We calculate $\Upsilon$ for every perturbation point and for all directions in the span of $\Delta\phi=$0--180$^\circ$ relative to $\mathbf{n}_{p1}$ (Fig.~\ref{stress_vel_corr_GL_allP}). For normal stress perturbations relatively more parallel to $\mathbf{n}_{p1}$ (closer to 0$^\circ$ or 180$^\circ$), there is a much stronger correlation between changes in normal stress and changes in ice surface speed (and hence the ice flux) than for normal stress perturbations that are relatively more parallel to $\mathbf{n}_{p2}$ (closer to 90$^\circ$).  
\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/stress_vel_corr_GL_allP.pdf}
    \caption{Correlation between the change in normal stress and the change in ice surface speed along the GL (i.e., $\Upsilon$ from (\ref{upsilon})). The horizontal axis shows how $\Upsilon$ varies as a function of the direction of the normal stress, rotated counterclockwise from $\mathbf{n}_{p1}$. Colored, dotted curves each represent a single perturbation experiment and the thick black curve is their mean for a given value of $\Delta\phi$. \textbf{SP: We should swap the order of this figure with the next one, since this is the 'continuous' version of that one (see related note in caption below).}}
	\label{stress_vel_corr_GL_allP}
\end{figure}
By plotting spatial maps of $\Upsilon$ for every perturbation point of interest in the domain, we show that there is generally a stronger correlation, in general, between changes in normal stress and velocity along $\mathbf{n}_{p1}$ (Fig.~\ref{stressDiff_velDiff_GL_allPerturb}a) than along $\mathbf{n}_{p2}$ (Fig.~\ref{stressDiff_velDiff_GL_allPerturb}b).  

>>>>

%\usepackage{appendix}
%\usepackage{xcite}
%\usepackage{xr}
%\usepackage{hyperref}
%\makeatletter
%\newcommand*{\addFileDependency}[1]{% argument=file name and extension
%  \typeout{(#1)}
%  \@addtofilelist{#1}
%  \IfFileExists{#1}{}{\typeout{No file #1.}}
%}
%\makeatother

%\newcommand*{\myexternaldocument}[1]{%
%    \externaldocument{#1}%
%    \addFileDependency{#1.tex}%
%    \addFileDependency{#1.aux}%
%}

%\myexternaldocument{supplement}
%\listfiles

@@

\begin{document}
\title{Diagnosing the sensitivity of grounding line flux to changes in sub-ice shelf melting}

@@

\begin{figure}
\centering
\includegraphics[width=0.4\linewidth]{./figs/mismip_larsenc.pdf}
\caption{Plan view of steady-state surface speed for MISMIP+ a) and Larsen C Ice Shelf b). The black curves show the grounding lines.}
\label{mismip_larsenc}
\end{figure}

@@

Our experiments are conducted in a manner similar to those of \citet{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over grid cells or square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, we use a uniform, hexagon-type, around 2-km mesh (the cell is not regular hexagon; for simplicity we use the term 2-km in the following). For the Larsen C Ice Shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we perturb the thickness of the cell (hexagon) in the mesh. For Larsen C, we only use 20$\times$20 km square boxes \citep[the same resolution as in][]{reese2018}. Lastly, for the MISMIP+ 2-km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is laterally buttressed).  We also perturb only half of the ice shelf, $y>40$ km, due to the symmetry about the center line, though we analyze the response to these perturbations over the entire model domain. The horizontal coordinates of model domains can be found in Fig.~\ref{mismip_larsenc}.

Similarly to \citet{reese2018}, we define a GLF ``response number'',
\begin{equation}
N_r = \left(\frac{R}{P}\right)^k,
\end{equation}
where $R$ is the ice flux change integrated along the entire grounding line, $P$ is the mass change associated with a single cell or grid-box perturbation and $k$ is a power-law index that allows for the possibility of a nonlinear relationship between ice shelf buttressing and the change in GLF \citep[see also][]{schoof2007}. Here, we use $k=1/n$ with $n=3$.

%Despite the existence of many possible factors linking GLF to ice shelf properties (ice flow direction, horizontal gradients in ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc.), here we mainly examine model stress fields and the distance between perturbation points and the GL. This is because, as we show below, these factors correlate closely with the sensitivity of changes in GLF to the imposed ice shelf perturbations. 
The metric that is suspicious of connecting GLF and basal melting is the buttressing of ice shelf, which generally resists ice flow across grounding line.
%\textbf{SP: Maybe we can skip the previous sentences and just skip to some form of the next one? We can motivate this simply by saying that the reason we use a buttressing number like Furst et al is that it's a popular/standard and easy way to connect local ice shelf perturbations to the changes in GLF from the Reese experiments.} 
\textbf{Tong: Agreed. The old sentence is unnecessarily redundant. I change it to a new, short sentence}To incorporate the local stress field and its buttressing capacity into our analysis, we also calculate a ``buttressing number'', $N_b$, analogous to that from \citet{furst2016},
\begin{equation}
N_b\left(\mathbf{n}\right)=1-\frac{\mathbf{n}\cdot\mathbf{\sigma}\cdot\mathbf{n}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=\frac{1}{2}\left(1-{\rho_i}/{\rho_w}\right)gH$) and $\rho_i$ (910\,kg\,m$^{-3}$) and $\rho_w$ (1028\,kg\,m$^{-3}$) are the density of ice and ocean water, respectively. $\sigma_{nn}=\mathbf{n}\cdot\mathbf{\sigma}\cdot\mathbf{n}$, is the normal stress along a specified horizontal direction given by the normal vector $\mathbf{n}$ (see details in below).

@@

A decrease in ice shelf buttressing tends to lead to an increase in GLF \citep[e.g.,][]{gagliardini2010} and intuitively we expect that the GLF should be relatively more sensitive to melt changes that occur in regions of relatively larger buttressing. Here, we aim to better understand and quantify the relationship between the local ice shelf buttressing ``strength'' (given by $N_b$) and changes in GLF (given by $N_r$). Can the change in GLF for a given melt perturbation be quantified (in the predictive sense) as a function of the buttressing number, i.e., a linear regression relationship $N_r\left(N_b\right)$?

In Fig.~\ref{mismip_Nb_GLF_regression}, we show correlations between $N_b$ and $N_r$ for perturbations applied to the MISMIP+ domain when $N_b$ from Eq.~(\ref{butN}) is defined based on the direction corresponding to the first principal stress ($\mathbf{n}_{p1}$; Fig.~\ref{mismip_Nb_GLF_regression}b), the second principle stress ($\mathbf{n}_{p2}$; Fig.~\ref{mismip_Nb_GLF_regression}c) and ice flow ($\mathbf{n}_{f}$; Fig.~\ref{mismip_Nb_GLF_regression}d), respectively. For these linear regression analysis, we remove from the comparison points that are 1) weakly buttressed ($x>480$\,km, where the ice shelf starts to become unconfined) and 2) the magnitude of the thickness gradient $\left|\nabla H\right|>7\times 10^{-3}$ (an arbitrary choice to remove perturbation points close to grounding line; Fig.~\ref{mismip_Nb_GLF_regression}a). Analogous results for including all perturbation points can be seen in Fig.~\ref{mismip_Nb_GLF_regression_all_points}. Clearly, a much stronger $N_r\left(N_b\right)$ linear regression emerges when we remove perturbation points that are close to grounding line and to the unbuttressed calving front. In addition, we find a relatively better linear regression function when we calculate $N_b$ along $\mathbf{n}_{p1}$ than along $\mathbf{n}_{p2}$ and $\mathbf{n}_{f}$. We discuss possible reasons for this further below.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{./figs/mismip_Nb_GLF_regression.pdf}
    \caption{(a) Blue dots are the selected perturbation points for $N_r\left(N_b\right)$ linear regression analysis and the black dots represent grid cells along the grounding line. (b) Modeled $N_r$ from perturbation experiments versus fitted $N_r$ as a function of $N_b$ that is calculated along $\textbf{n}_{p1}$; (c) Modeled $N_r$ from perturbation experiments versus fitted $N_r$ as a function of $N_b$ that is calculated along $\textbf{n}_{p2}$; (d) Modeled $N_r$ from perturbation experiments versus fitted $N_r$ as a function of $N_b$ that is calculated along $\textbf{n}_{f}$. $r$ is the correlation coefficient. \textbf{Tong: Do we need to include the information of significant level along with correlation coefficient?}}
\label{mismip_Nb_GLF_regression}
\end{figure}

%\subsection{The buttressing directions}
%\subsection{Correlation dependence on buttressing direction}
\subsection{Directional dependence of buttressing}\label{directional_buttressing}

According to Eq. (\ref{butN}), the buttressing number at any perturbation point varies depending on the chosen direction. \citet{furst2016} calculated $N_b$ along two directions, the ice flow direction ($\mathbf{n}_f$) and the direction corresponding to the second principal stress ($\sigma_{p2}$) ($\mathbf{n}_{p2}$). They found that the latter---the direction corresponding to the maximum compressive stress (or the least extensional stress)---best identified ``passive'' regions, where the ice shelf provides little or no buttressing. In Fig.~\ref{mismip_r2_all_direction}a, we plot the correlation coefficients ($r$) for the linear regression function $N_r\left(N_b\right)$ where the direction ($\textbf{n}$) corresponding to $\sigma_{nn}$ varies continuously between 0 and 180$^\circ$ (as a function of the angle $\Delta \phi$) relative to $\mathbf{n}_{p1}$. We find the largest correlation coefficient ($r=0.964$) when $N_b$ is aligned with $\mathbf{n}_{p1}$ ($\Delta \phi=0^\circ$) and the smallest correlation coefficient ($r=0.393$) when $N_b$ is aligned with $\mathbf{n}_{p2}$ ($\Delta \phi\approx 90^\circ$). We can also see the corresponding variations in buttressing number $N_b$ in Fig.~\ref{Nb_Deltaphi}, where the maximum buttressing number is calculated along $\mathbf{n}_{p2}$ according to Eq. (\ref{butN}).
% for different directions with respect to the first principal stress and flow direction, respectively (Fig \ref{mismip_r2_all_direction}). Differently, we find that $N_b$ along the $\sigma_{p1}$ direction shows the best regression performance, whereas the $\sigma_{p2}$ appears to show the weakest correlations to $N_r$. This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$).

@@

A similar conclusion can be reached when the buttressing number is calculated by projecting $\sigma_{nn}$ onto $\mathbf{n}_{f}$\footnote{In \citet{furst2016}, the ice flow direction is the second direction considered when calculating the local ice shelf buttressing number (with the first direction considered being $\mathbf{n}_{p2}$).} (Fig.~\ref{mismip_r2_all_direction}b).
%the blue line shows the $N_b$:$N_r$ correlation for the case of rotating relative to $\mathbf{n}_{f}$. %(the flow direction experiments are analyzed independently from the $\sigma_{p1}$ direction experiments, and can therefore be used as an approach for verification). 
For around 50\% of the perturbation points, the angles of flow directions are 30--50$^\circ$ ($\Delta\varphi$) more than that of $\mathbf{n}_{p1}$ (from Fig.~\ref{mismip_r2_all_direction}c), which is consistent with the phase difference between the blue and red curves in Fig.~\ref{mismip_r2_all_direction}a and b. For example, for the $\mathbf{n}_{f}$ curve (blue), we have the smallest value at $\Delta\phi\approx$ 38$^\circ$, i.e., 38$^\circ$ relative to the flow direction $\mathbf{n}_{f}$. Since the angles of $\mathbf{n}_{f}$ are 30--50$^\circ$ ($\Delta\varphi$) more than that of $\mathbf{n}_{p1}$, the lowest value should occur at 68--88$^\circ$ ($\Delta\phi+\Delta\varphi$) relative to $\mathbf{n}_{p1}$, which is corresponding to $\mathbf{n}_{p2}$. This is in consistent with the position of the smallest value for the $\sigma_{p1}$ direction curve (red; $\Delta\phi\approx$ 90$^\circ$). %\textbf{(SFP: See comments in source text regarding this paragraphy.)}

@@

Based on the discussion above and the work of \citet{furst2016}, it appears that $N_b(\mathbf{n}_{p2})$ provides a good indicator for the maximum buttressing at any location on the shelf and for identifying local passive ice shelf areas. Yet it may not be a good metric for quantifying the sensitivity of GLF response to sub-shelf basal melt perturbations. In the following sections, we explain how and why changes in GLF are sensitive to the choice of direction when calculating the ice shelf buttressing number.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{./figs/mismip_r2_all_direction.pdf}
    \caption{(a) Correlation coefficients for the linear regression function $N_r\left(N_b\right)$ where $\textbf{n}$ is rotated counterclockwise by $\Delta\phi$ degrees relative to $\mathbf{n}_{p1}$; (b) Correlation coefficients for the linear regression function $N_r\left(N_b\right)$ where $\textbf{n}$ is rotated counterclockwise by $\Delta\phi$ degrees relative to $\mathbf{n}_{f}$; (c) Histogram of the angular differences between $\mathbf{n}_{f}$ and $\mathbf{n}_{p1}$ ($\Delta\varphi$). The perturbation points analyzed here are the same as in Fig.~\ref{mismip_Nb_GLF_regression}a.}
    \label{mismip_r2_all_direction}
\end{figure}

%\subsection{Possible controlling factors}
%\subsection{Understanding the dependence on buttressing direction}
%\subsection{Direction of velocity changes around perturbation points}
%\subsection{Ice speed changes in the vicinity of perturbation points}
\subsection{Directional dependence of ice speed changes}

The correlation between changes in GLF and normal stress (buttressing) directions can be partly understood by examining the spatial patterns of ice speed change associated with thickness perturbations. Below we discuss the ice speed changes in the vicinity of perturbation points and along grounding line, aiming to seek the patterns of how the perturbations propagate to upstream.  %Further clues of the impacts from different chosen directions can be found in Fig.~\ref{perturb_theta_velchange}.

In Fig.~\ref{perturb_theta_velchange}, we plot histograms of the maximum (red) and minimum (blue) speed changes around each perturbation point as a function of angular distance relative to $\mathbf{n}_{p1}$ ($\Delta\phi$, as in Fig.~\ref{mismip_r2_all_direction}a). That is, we count the number of neighboring cells that have maximum (most increased or least decreased) and minimum (most decreased or least increased) speed changes (for each cell, there is only one neighbor with maximum or minimum speed change).  %Figure \ref{perturb_theta_velchange} contains only the points from Fig.~\ref{mismip_Nb_GLF_regression}d (i.e., filtered according to ice thickness gradient). 
The maximum speed changes cluster within 0 to 45$^\circ$ and the minimum speed changes cluster within 60 to 90$^\circ$ relative to $\mathbf{n}_{p1}$ , indicating that the majority of ice speed increases in the immediate region of a perturbation are more closely aligned with $\mathbf{n}_{p1}$ (and likewise, the majority of velocity decreases are more closely aligned with $\mathbf{n}_{p2}$). This finding supports the hypothesis that the thickness perturbations induce velocity increases along favored directions and that, for the MISMIP+ domain, $\mathbf{n}_{p1}$ is favored. This is the pattern of ice speed changes around perturbations, the start points of perturbation propagation. What is the case along grounding line, the upstream regions we are most interested in?

%\textbf{SP: I think it would help if we try to connect this more clearly to the next section. That is, here we show that a local perturbation leads to speed increases along preferred directions. Below, we show that local perturbations lead to similar speed increases (also along preferred directions) but at the GL.}

@@

\begin{figure}
	\centering
    \includegraphics[width=0.8\linewidth]{figs/perturb_theta_velchange.pdf}
    \caption{Histograms for the maximum (red) and minimum (blue) percent speed change in grid cells adjacent to a thickness perturbation point, plotted as a function of angular distance with respect to $\mathbf{n}_{p1}$. Points analyzed are those from Fig.~\ref{mismip_Nb_GLF_regression}d.}
	\label{perturb_theta_velchange}
\end{figure}

%\subsection{The changes of velocity and stress along GL}
%\subsection{Speed and stress changes along the GL}

Therefore, further insight into the importance of $\mathbf{n}_{p1}$ is given by examining changes in the state of stress and velocity along grounding line as a result of a perturbation in local ice thickness. To this end, we define the metric $\Upsilon$,  
\begin{equation}
    \Upsilon= \textrm{Corr}\left(\frac{\overrightarrow{\sigma_{p}}-\overrightarrow{\sigma_{c}}}{\overrightarrow{\sigma_{c}}}, \frac{\overrightarrow{u_{p}}-\overrightarrow{u_{c}}}{\overrightarrow{u_{c}}}\right),
\label{upsilon}
\end{equation}
%\begin{equation}
%    \Psi= \frac{\sigma_{p}-\sigma_{c}}{\sigma_{c}}-\frac{u_{p}-u_{c}}{u_{c}},
%\label{psi}
%\end{equation}
%\textbf{SP: need to update this equation to clarify the following things: 1) upsilon is scalar function of x,y (that is, we calculate a single value of it for each perturbation point on the ice shelf), 2) the two terms in parentheses are vectors of points along the GL, 3) the operation here is a dot product, so that the vectors collapse to a scaler value, which is what is evalutated).}
where the subscripts $p$ and $c$ denote the perturbation experiments and the ``control'' (i.e., the initial condition), respectively. $\overrightarrow{\sigma}$ and $\overrightarrow{u}$ denote the vectors of stress component and ice speed for all grid cells along grounding line, respectively. $\Upsilon$, a correlation coefficient, is a measure of the consistency between the magnitude and sign of the change in the model stress and velocity states along grounding line between the control and perturbation experiments.

By plotting spatial maps of $\Upsilon$ for every perturbation point of interest in the domain, we show that there is generally a stronger correlation, in general, between changes in normal stress and ice speed along $\mathbf{n}_{p1}$ (Fig.~\ref{stressDiff_velDiff_GL_allPerturb}a) than along $\mathbf{n}_{p2}$ (Fig.~\ref{stressDiff_velDiff_GL_allPerturb}b). Further, we calculate $\Upsilon$ for every perturbation point and for all directions in the span of $\Delta\phi=$0--180$^\circ$ relative to $\mathbf{n}_{p1}$ (Fig.~\ref{stress_vel_corr_GL_allP}). Clearly, in a continuous manner, we confirm that there is a much stronger correlation between changes in normal stress calculated along $\mathbf{n}$ more parallel to $\mathbf{n}_{p1}$ (closer to 0$^\circ$ or 180$^\circ$) and changes in ice surface speed (and hence the ice flux) than along $\mathbf{n}$ that are relatively more parallel to $\mathbf{n}_{p2}$ (closer to 90$^\circ$).

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{./figs/stressDiff_velDiff_GL_allPerturb.pdf}
    \caption{Spatial distribution of the correlation coefficient $\Upsilon$ from Eq.~(\ref{upsilon}) over the MISMIP+ domain for normal stress changes calculated parallel to $\mathbf{n}_{p1}$ (a) and $\mathbf{n}_{p2}$ (b). Note that $\Upsilon$ is a metric indicating the changes of stress and ice speed along grounding line. The thick dashed black line is grounding line.}
	\label{stressDiff_velDiff_GL_allPerturb}
\end{figure}

\begin{figure}
	\centering
    \includegraphics[width=0.8\linewidth]{./figs/stress_vel_corr_GL_allP.pdf}
    \caption{Correlation between the change in normal stress and the change in ice surface speed along grounding line (i.e., $\Upsilon$ from Eq.~(\ref{upsilon})). The horizontal axis shows how $\Upsilon$ varies as a function of the direction of the normal stress $\mathbf{n}_{p1}$, rotated counterclockwise from $\mathbf{n}_{p1}$. The blue shaded area is the range of all single perturbation experiments and the thick black curve is their mean value.}
	\label{stress_vel_corr_GL_allP}
\end{figure}
